The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Finish chapter 4.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@633 4525493e-7705-40b1-a816-d608a930855b

The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin was responsible for invoking
data-service APIs and parsing retrieved data into the DAS data format
suitable for storage, JSON. Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
we dealt with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values as regular expression patterns into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used {\it logical\_file\_name} notation to identify the name of the
logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key {\it file}.\footnote{We
also deal with physical file names, but still used {\it file} to refer to them in
DAS notations, since simple regular expressions were used to identify the
meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type {\it file} in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. {\it file=/abc*}.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. For example,
in aforementioned example, using {\it file} in DAS query cause invokation of DBS 
{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
This approach gave power to end-users to express the meaning of their queries, especially
when they deal with numbers and achieve accuracy of results from DAS. 
For instance, users were able to type {\it run > 10 and run < 100}
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full free text based keyword search is under development within MongoDB.

The DAS Analytics DB played a special role. It collected information
about user requests placed to the system. Each request was recorded. Upon its
decomposition into set of selection keys and conditions we also recorded
API name and input parameters into the Analytics database and 
keep updating their counter information for repeated calls. 
This allows to keep track of frequency of API calls and establish 
pre-fetch strategies for most common queries.

The DAS caching system was used to store data-service results and aggregated
information in form of DAS records. As it shown in figure \ref{DAS_cache} 
the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
and aggregate data over there. The independent set of daemons, DAS robots, were
constantly monitor Analytics DB for most popular requests and pre-fetch most
common data into DAS cache.

@@

\subsection{Data-services and aggregation}
DAS provided an interface to add new CMS data-services via data-service plugins.
Each plugin consists of its DAS map and data parser suitable for data-service output
format(s).
The data-service DAS map contains information about data-service APIs,
input and output parameters, data-service
format and their mapping to DAS keys as well as relation to other sub-systems. 
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
patterns, evaluated by python during run time
execution, to validate provided set of input paremters and DAS keys.}:
\begin{verbatim}
{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", 
              "das_key": "site"}], 
 "_id": "4ae1f8dfe2194e4581000014", 
 "system": "sitedb", 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", 
              "map": "site.name"}]}
\end{verbatim}
This information was stored into DAS Mapping DB.

The DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS and their
relationships. Based information stored there we were able to retrieve appropriate data
from different data-services, re-map them into DAS notations and
aggregate them on demand.
For instance, the Data Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS data files. In former case, DBS holds information about
files and their relationship to other physics objects, such as run and luminosity, 
while PhEDEx stores information about the file location, such as sites which
holds copy of the files and details about site storage elements.
Upon a user request to find a file, we were able to identify appropriate set of
DBS and PhEDEx APIs, look-up data, re-format their output to DAS notations and
store results into DAS cache.
Once data were retreived from data-services an additional step to aggregate
matched objects in DAS cache were applied. For instance, both data-service outputs 
contained a file name in them. We use it as a common key to merge records in DAS cache.
This merged (aggregated) results represent DAS records. Each DAS
record contains a standard DAS header, which identify look-up time,
request url, input parameters and API method as well as response version, expiration
time and checksum of the output. Here is an example of individual DAS record:
\begin{verbatim}
{"_id": "4ae1e7ffe2194e4243000003", 
 "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
 "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 1.2158100605010986, 
          "api": ["CMStoSiteName"], 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
        ]
}
\end{verbatim}
This DAS record represents aggregated information between two SiteDB APIs, 
{\it CMStoSAMName} and {\it CMStoSiteName}, where API results were merged 
under the {\it site} key. The information provided by {\it das} key shows 
DAS header, which contains information about API name, location, execution 
time and lifetime of the data.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc.

We used RESTful model \cite{REST} for all DAS cache server APIs.
It allows us to share core library and other components during development of 
various DAS clients, e.g. CLI, web interface, robots, etc. and hides
complexity of access to individual data-services and their security policies
as well as proper organize caching solution.
For instance, to access on-line Run Summary information we were forced to use
Grid certificates. Instead of propagating user certificate to data-service
the DAS cache server works as a proxy to this server and used host certficiate 
to serve such requests on behalf of the users.

>>>>

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map,
data parser(s) suitable for data-service output format(s) into DAS common
data format (JSON) and was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
patterns, evaluated during run time, to validate provided set of input 
paremters and DAS keys.}:
\begin{verbatim}
{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", 
              "das_key": "site"}], 
 "_id": "4ae1f8dfe2194e4581000014", 
 "system": "sitedb", 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", 
              "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
we dealt with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS Mapping DB}
DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values as regular expression patterns into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used {\it logical\_file\_name} notation to identify the name of the
logical file, while the same entity was called {\it lfn} in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key {\it file}.\footnote{We
also deal with physical file names, but still used {\it file} to refer to them in
DAS notations, since simple regular expressions were used to identify the
meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. {\it file}, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type {\it file} in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. {\it file=abc*}.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. For example,
in aforementioned example, using {\it file} in DAS query cause call of DBS 
{\it listFiles} and PhEDEx {\it fileReplicas} APIs, respectively.
This approach gave power to end-users to express the meaning of their queries, especially
when they deal with numbers and achieve accuracy of results from DAS. 
For instance, users were able to type $run>10\,\,and\,\, run < 100$
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full free text based keyword search is under development within MongoDB.

\subsection{DAS Analytics DB}
The DAS Analytics DB played a special role. It collected information
about user requests placed to the system. Each request was recorded. Upon its
decomposition into set of selection keys and conditions we also recorded
API name and input parameters into the Analytics database and 
keep updating their counter information for repeated calls. 
This allows to keep track of frequency of API calls and establish 
pre-fetch strategies for most common queries.

\subsection{DAS caching system}
The DAS caching system was used to store data-service results and aggregated
information in form of DAS records. As it shown in figure \ref{DAS_cache} 
the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
and aggregate data over there. The independent set of daemons, DAS robots, were
constantly monitor Analytics DB for most popular requests and pre-fetch most
common data into DAS cache.

@@

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS data files. In former case, DBS 
stored information about files and their relationship to other 
physics objects, such as run and luminosity, while PhEDEx kept
information about the file location, e.g. details about sites 
holding copy of the files. Upon a user request to find a file, 
we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects. For instance, since both data-services returns in their
output a file name we used as a common key during aggregation.
This merged results represented a set of DAS records. Each DAS
record contains a standard DAS header, which identify look-up time,
request url, input parameters and API method as well as response 
version, expiration time and checksum of the output. Usage of this 
information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc. Here is an example of DAS record representing
aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
and {\it CMStoSiteName}, with API results merged under the {\it site} key.
\begin{verbatim}
{"_id": "4ae1e7ffe2194e4243000003", 
 "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
 "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 1.2158100605010986, 
          "api": ["CMStoSiteName"], 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
        ]
}
\end{verbatim}

%We used RESTful model \cite{REST} for all DAS cache server APIs.
%It allows us to share core library and other components during development of 
%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
%complexity of access to individual data-services and their security policies
%as well as proper organize caching solution.
%For instance, to access on-line Run Summary information we were forced to use
%Grid certificates. Instead of propagating user certificate to data-service
%the DAS cache server works as a proxy to this server and used host certficiate 
%to serve such requests on behalf of the users.
